---
title: "Final Project"
author: "Shraddha Hemant Kadam (sxk190069@utdallas.edu)"
date: "08/04/2020"
output:
  pdf_document:
    latex_engine: xelatex
  latex_engine: default'
---


```{r package loading, warning=FALSE}

if(!require("pacman")) install.packages("pacman")
pacman::p_load(forecast, leaps, ggcorrplot, corrplot, tidyverse, reshape, gplot, ggmap, mlbench, data.table, factoextra)
search()
theme_set(theme_classic())
install.packages("funModeling")
install.packages("flexclust")
install.packages("tidyverse")
install.packages("Hmisc")
library(gridExtra)
install.packages("RCurl")
install.packages("skimr")
install.packages("grid")
library(flexclust)
install.packages("modeltools")
install.packages("stats4")
library(stats4)
library(modeltools)
library(funModeling)
library(skimr)
library(grid)
library(RColorBrewer)
library(RCurl)
library(dplyr)
library(ggplot2)
library(reshape2) 
library(gridExtra) # allow us to plot two plots in a row
options(scipen=999) # disable scientific notation
pacman::p_load(caret, corrplot, glmnet, tidyverse, data.table, gplots, ggrepel)
```

```{r read datasets, warning= FALSE}
#loading updated datasets from github
supply_kcal_data <- read.csv(text = getURL("https://raw.githubusercontent.com/mariarencode/COVID_19_Dataset_Challenge/master/Cleaned_Datasets/Food_Supply_kcal_Data.csv"))
supply_kg_data <- read.csv(text = getURL("https://raw.githubusercontent.com/mariarencode/COVID_19_Dataset_Challenge/master/Cleaned_Datasets/Food_Supply_Quantity_kg_Data.csv"))
protein_data <- read.csv(text = getURL("https://raw.githubusercontent.com/mariarencode/COVID_19_Dataset_Challenge/master/Cleaned_Datasets/Protein_Supply_Quantity_Data.csv"))
fat_data <- read.csv(text= getURL("https://raw.githubusercontent.com/mariarencode/COVID_19_Dataset_Challenge/master/Cleaned_Datasets/Fat_Supply_Quantity_Data.csv"))
```

```{r EDA, warning= FALSE}
#display head and tail of datasets

head(fat_data,5)
tail(fat_data,5)
head(supply_kcal_data,5)
tail(supply_kcal_data,5)
head(supply_kg_data,5)
tail(supply_kg_data,5)
head(protein_data,5)
tail(protein_data,5)

#get summary of datasets
summary(fat_data)
summary(supply_kcal_data)
summary(supply_kg_data)
summary(protein_data)


df_status(fat_data)
df_status(supply_kcal_data)
df_status(supply_kg_data)
df_status(protein_data)

#handeling missing values
df_status(fat_data, print_results = FALSE) %>% select(variable, q_na, p_na) %>% arrange(-q_na)
df_fat_data<- fat_data
df_fat_data[is.na(df_fat_data)]=0


df_supply_kcal_data <- supply_kcal_data
df_supply_kcal_data[is.na(df_supply_kcal_data)]=0

df_supply_kg_data<- supply_kg_data
df_supply_kg_data[is.na(df_supply_kg_data)]=0

df_protein_data<- protein_data
df_protein_data[is.na(df_protein_data)]=0


#check to see if there are missing values in datasets
sum(is.na(df_fat_data))
sum(is.na(df_supply_kcal_data))
sum(is.na(df_protein_data))
sum(is.na(df_supply_kg_data))

#checking the number of zeros
df_status(df_fat_data, print_results = FALSE) %>% select(variable, q_zeros, p_zeros) %>% arrange(-q_zeros)

df_status(df_supply_kcal_data, print_results = FALSE) %>% select(variable, q_zeros, p_zeros) %>% arrange(-q_zeros)

df_status(df_supply_kg_data, print_results = FALSE) %>% select(variable, q_zeros, p_zeros) %>% arrange(-q_zeros)

df_status(df_protein_data, print_results = FALSE) %>% select(variable, q_zeros, p_zeros) %>% arrange(-q_zeros)


#get detailed summary by providing larger set of statistics
skim(df_fat_data)
skim(df_supply_kcal_data)
skim(df_supply_kg_data)
skim(df_protein_data)

```

```{r plots, warning= FALSE}
#scatterplot
plot(df_fat_data[, c("Active","Confirmed", "Deaths", "Recovered")], col= "green")


ggplot(df_fat_data) +
  geom_boxplot(aes(x = Active, y = Population), 
               fill = "gold1", outlier.color = "firebrick2") + 
  xlab("Active") + ggtitle("COVID-19 Active cases")

ggplot(df_fat_data) +
  geom_boxplot(aes(x = Confirmed, y = Population), 
               fill = "gold1", outlier.color = "firebrick2") + 
  xlab("Active") + ggtitle("COVID-19 Confirmed cases")

ggplot(df_fat_data) +
  geom_boxplot(aes(x = Deaths, y = Population), 
               fill = "gold1", outlier.color = "firebrick2") + 
  xlab("Active") + ggtitle("COVID-19 Death cases")

ggplot(df_fat_data) +
  geom_boxplot(aes(x = Recovered, y = Population), 
               fill = "gold1", outlier.color = "firebrick2") + 
  xlab("Active") + ggtitle("COVID-19 Recovered cases")
```


```{r graph pie chart, warning= FALSE}
pie_charts<- function(data, title_text) {
  chart_data <- data.frame(
    food_groups = names(data[2:24]),
    food_percentage = as.numeric(data[2:24])
  )
  
  ggplot(chart_data, aes(x="", y=food_percentage, fill=food_groups)) +
    geom_bar(stat="identity", width=1) +
    theme_classic() +
    theme(plot.title = element_text(hjust=0.5),
          axis.line = element_blank(),
          axis.text = element_blank(),
          axis.ticks = element_blank()) +
    labs(fill = "food_group",
         x = NULL,
         y = NULL,
         title = title_text) + 
    coord_polar("y")
}

# Graph bar plots
bar_plot <- function(data, title_text, rounding_digits){
  bar_plot_data <- data.frame(
    groups = names(data),
    percentage = as.numeric(data)
  )
  ggplot(data=bar_plot_data, aes(x=groups, y=percentage, fill=groups)) +
    geom_bar(stat="identity", position=position_dodge())+
    geom_text(aes(label=paste(round(percentage,rounding_digits), "%")), vjust=1.6, color="black",
              position = position_dodge(0.4), size=3.5)+
    scale_fill_brewer(palette="Paired")+
    theme_minimal() +
    labs(x = NULL,
         y = "Percentage out of Total Population",
         title = title_text)
}
```

```{r India_data_Analysis, warning= FALSE}
# India data analysis and plots
# Pull out India data
India_fat_data <- df_fat_data[which(df_fat_data$Country == "India"),]
India_supply_kcal_data <- df_supply_kcal_data[which(df_supply_kcal_data$Country == "India"),]
India_supply_kg_data <- df_supply_kg_data[which(df_supply_kg_data$Country == "India"),]
India_protein_data <- df_protein_data[which(df_protein_data$Country == "India"),]
# Pie charts for India 
India_fat_pie_chart <- pie_charts(India_fat_data, "Percentage of Fat Consumed (India)")
India_food_kcal_pie_chart <- pie_charts(India_supply_kcal_data, "Percentage of Energy Consumed (India)")
India_food_kg_pie_chart <- pie_charts(India_supply_kg_data, "Percentage of Food Supply (India)")
India_protein_pie_chart <- pie_charts(India_protein_data, "Percentage of Protein Consumed (India)")
# Bar plots for US 
India_obesity_undernourished_bar_plot <- bar_plot(India_supply_kg_data[25:26], "Obesity and Undernourished Percentages (India)", 3)
India_covid_cases_bar_plot <- bar_plot(India_supply_kg_data[27:30], "COVID-19 Cases Percentages (India)", 3)
```


```{r US_data_analysis, warning= FALSE}
# US data analysis and plots
# Pull out US data
us_fat_data <- df_fat_data[which(df_fat_data$Country == "United States of America"),]
us_supply_kcal_data <- df_supply_kcal_data[which(df_supply_kcal_data$Country == "United States of America"),]
us_supply_kg_data <- df_supply_kg_data[which(df_supply_kg_data$Country == "United States of America"),]
us_protein_data <- df_protein_data[which(df_protein_data$Country == "United States of America"),]
# Pie charts for US 
us_fat_pie_chart <- pie_charts(us_fat_data, "Percentage of Fat Consumed (US)")
us_food_kcal_pie_chart <- pie_charts(us_supply_kcal_data, "Percentage of Energy Consumed (US)")
us_food_kg_pie_chart <- pie_charts(us_supply_kg_data, "Percentage of Food Supply (US)")
us_protein_pie_chart <- pie_charts(us_protein_data, "Percentage of Protein Consumed (US)")
# Bar plots for US 
us_obesity_undernourished_bar_plot <- bar_plot(us_supply_kg_data[25:26], "Obesity and Undernourished Percentages (US)", 3)
us_covid_cases_bar_plot <- bar_plot(us_supply_kg_data[27:30], "COVID-19 Cases Percentages (US)", 3)
```

```{r fat_Consumption, warning= FALSE}
# Fat consumption from food groups comparison
grid.arrange(us_fat_pie_chart, India_fat_pie_chart, nrow = 2)
```
From the above pir charts, it seems like India consumes more fat from Alcoholic Beverages than the US and US consumes more fats from pulses/Spices than that consumed in India. But overall, it seems like the difference in each category isn't drastically different between the two countries.

```{r Energy_Consumption, warning= FALSE}
grid.arrange(us_food_kcal_pie_chart, India_food_kcal_pie_chart, nrow = 2)
```
For the energy consumed(in kilo calories), India has slightly higher percentage in the Animal product category whereas US is higher in percentage in Treenuts and Sugar Crops.

```{r supply, warning= FALSE}
grid.arrange(us_food_kg_pie_chart, India_food_kg_pie_chart, nrow = 2)

```
Looking at the food supply comparision, we see that India has more supply of Animal Products than US whereas US has slightly highe supply of Fruits excluding wine than India.

```{r protein consumption, warning= FALSE}
grid.arrange(us_protein_pie_chart, India_protein_pie_chart, nrow = 2)

```
From the above charts we can see that India's protein consumption from Alcoholic Beverages,Pulses /Spices /Starchy roots and Fish Seafood in more than US. On the other side US's protein consumption is more from Fruits and Animal Fats/ Animal Products.

```{r obesity, warning= FALSE}
grid.arrange(us_obesity_undernourished_bar_plot,India_obesity_undernourished_bar_plot, nrow = 2)

```
Looking at the obesity and undernourished rate bar plot above, US has higher obesity percentage whereas India has higher Undernourished percentage.

```{r covid, warning= FALSE}
grid.arrange(us_covid_cases_bar_plot, India_covid_cases_bar_plot, nrow = 2)

```
And lastly, taking a quick look at the percentage of COVID cases. US confirmed cases is 1.415% of the entire population, whereas that of India is 0.129% of the entire population. The recovery rate of India is higher than that of US.



```{r pricipal component analysis, warning= FALSE}
pcs1<- prcomp(na.omit(df_protein_data[,-c(1,25,26,32)]), scale. = T)
summary(pcs1)  

pcs2 <- prcomp(na.omit(df_supply_kg_data[,-c(1,25,26,32)]), scale. = T)
summary(pcs2)

pcs3_carbs <- prcomp(na.omit(df_supply_kcal_data[,-c(1,25,26,32)]), scale. = T)
summary(pcs3_carbs)  

pcs4_fats <- prcomp(na.omit(df_fat_data[,-c(1,25,26,32)]), scale. = T)
summary(pcs4_fats)  
```
since 99% is covered till pc24, we will consider variables till pc24

```{r Data Partition, warning= FALSE}
set.seed(42)
#randomly order the dataset 
rows_p<- sample(nrow(df_protein_data))
df_protein_data <- df_protein_data[rows_p,]

rows_c<- sample(nrow(df_supply_kcal_data))
df_supply_kcal_data<- df_supply_kcal_data[rows_c,]

rows_f<- sample(nrow(df_fat_data))
df_fat_data <- df_fat_data[rows_f,]

rows_k<- sample(nrow(df_supply_kg_data))
df_supply_kg_data<- df_supply_kg_data[rows_k,]

#splitting the data
set.seed(42)
split <- round(nrow(df_protein_data) * 0.8)
trainp.df <- df_protein_data[1:split, -c(1,25,26,28:32)]
testp.df <- df_protein_data[(split+1):nrow(df_protein_data), -c(1,25,26,28:32) ]

split <- round(nrow(df_supply_kcal_data) * 0.8)
trainc.df <- df_supply_kcal_data[1:split, -c(1,25,26,28:32)]
testc.df <- df_supply_kcal_data[(split+1):nrow(df_supply_kcal_data), -c(1,25,26,28:32) ]

split <- round(nrow(df_fat_data) * 0.8)
trainf.df <- df_fat_data[1:split, -c(1,25,26,28:32)]
testf.df <- df_fat_data[(split+1):nrow(df_fat_data), -c(1,25,26,28:32) ]

split <- round(nrow(df_supply_kg_data) * 0.8)
trains.df <- df_supply_kg_data[1:split, -c(1,25,26,28:32)]
tests.df <- df_supply_kg_data[(split+1):nrow(df_supply_kg_data), -c(1,25,26,28:32) ]



```


```{r Clustering, warning=FALSE}
#Calculating Distance

# Set row names to the utilities column

row.names(df_protein_data) <- df_protein_data[,1]
# Remove the utility column (include only numeric variables in Cluster Analysis)
protein.df <- df_protein_data[,-c(1,25,26,28:32)]
# Compute Euclidean distance
d <- dist(protein.df, method = "euclidean")
# Normalize and calculate distances
protein.df.norm <- sapply(protein.df, scale)
# Add row names: utilities
row.names(protein.df.norm) <- row.names(protein.df) 
# ALT: Compute normalized distance based on all 8 variables
d.norm <- dist(protein.df.norm, method = "euclidean")
d.norm

row.names(df_supply_kcal_data) <- df_supply_kcal_data[,1]
carbs.df <- df_supply_kcal_data[,-c(1,25,26,28:32)]
d <- dist(carbs.df, method = "euclidean")
carbs.df.norm <- sapply(carbs.df, scale)
row.names(carbs.df.norm) <- row.names(carbs.df) 
d.norm2 <- dist(carbs.df.norm, method = "euclidean")
d.norm2

row.names(df_fat_data) <- df_fat_data[,1]
fats.df <- df_fat_data[,-c(1,25,26,28:32)]
d <- dist(fats.df, method = "euclidean")
fats.df.norm <- sapply(fats.df, scale)
row.names(fats.df.norm) <- row.names(fats.df) 
d.norm3 <- dist(fats.df.norm, method = "euclidean")
d.norm3

row.names(df_supply_kg_data) <- df_supply_kg_data[,1]
supply.df <- df_supply_kg_data[,-c(1,25,26,28:32)]
d <- dist(supply.df, method = "euclidean")
supply.df.norm <- sapply(supply.df, scale)
row.names(supply.df.norm) <- row.names(supply.df) 
d.norm4 <- dist(supply.df.norm, method = "euclidean")
d.norm4


  ## Generate Dendrogram using multiple methods  
hc1 <- hclust(d.norm, method = "single")
plot(hc1, hang = -1, main = "Single Linkage", col = "cornflowerblue") +
  abline(h=2.7, lty = 1, lwd = 1, col = "firebrick4")

hc2 <- hclust(d.norm, method = "complete")
plot(hc2, hang = -1, main = "Complete Linkage", col = "dodgerblue") +
  abline(h=3.8, lty = 1, lwd = 1, col = "firebrick4")

hc3 <- hclust(d.norm, method = "average")
plot(hc3, hang = -1, main = "Average Linkage", col = "deepskyblue") +
  abline(h= 3.5, lty = 1, lwd = 1, col = "firebrick4")

hc22 <- hclust(d.norm2, method = "complete")
plot(hc22, hang = -1, main = "Complete Linkage", col = "dodgerblue") +
  abline(h=3.8, lty = 1, lwd = 1, col = "firebrick4")

hc32 <- hclust(d.norm3, method = "complete")
plot(hc32, hang = -1, main = "Complete Linkage", col = "dodgerblue") +
  abline(h=3.8, lty = 1, lwd = 1, col = "firebrick4")

hc42 <- hclust(d.norm4, method = "complete")
plot(hc42, hang = -1, main = "Complete Linkage", col = "dodgerblue") +
  abline(h=3.8, lty = 1, lwd = 1, col = "firebrick4")


## Check memberships in above clusters
memb1 <- cutree(hc1, k = 6)
memb1

memb2 <- cutree(hc2, k = 6)
memb2

memb3 <- cutree(hc3, k = 6)
memb3

memb22 <- cutree(hc2, k = 6)
memb22

memb32 <- cutree(hc32, k = 6)
memb32

memb42 <- cutree(hc42, k = 6)
memb42

#Validating clusters using heatmaps

row.names(protein.df.norm) <- paste(memb2, ": ", row.names(protein.df), sep = "")
# Heatmap 
heatmap(as.matrix(protein.df.norm), Colv = NA, hclustfun = hclust, col = brewer.pal(8, "PuBu"))

row.names(carbs.df.norm) <- paste(memb2, ": ", row.names(carbs.df), sep = "")
# Heatmap 
heatmap(as.matrix(carbs.df.norm), Colv = NA, hclustfun = hclust, col = brewer.pal(8, "PuBu"))

row.names(fats.df.norm) <- paste(memb2, ": ", row.names(fats.df), sep = "")
# Heatmap 
heatmap(as.matrix(fats.df.norm), Colv = NA, hclustfun = hclust, col = brewer.pal(8, "PuBu"))

row.names(supply.df.norm) <- paste(memb2, ": ", row.names(supply.df.norm), sep = "")
# Heatmap 
heatmap(as.matrix(supply.df.norm), Colv = NA, hclustfun = hclust, col = brewer.pal(8, "PuBu"))


#Assign the test observations to the clusters built upon the training set.
x <- as.kcca(hc2, protein.df.norm, k= 6)
clusterTrain1 <- predict(x)
clusterTest1 <- predict(x, newdata = testp.df)

a <- as.kcca(hc22, carbs.df.norm, k= 6)
clusterTrain2 <- predict(a)
clusterTest2 <- predict(a, newdata = testc.df)

b <- as.kcca(hc32, fats.df.norm, k= 6)
clusterTrain3 <- predict(b)
clusterTest3 <- predict(b, newdata = testf.df)

c <- as.kcca(hc42, supply.df.norm, k= 6)
clusterTrain4 <- predict(c)
clusterTest4 <- predict(c, newdata = tests.df)




#How many test-set observations were assigned to Clusters?
table(clusterTest1)
table(clusterTest2)
table(clusterTest3)
table(clusterTest4)

#Using the subset function, build data frames
proteintrain1 <- subset(trainp.df, clusterTrain1 == 1)
proteintrain2 <- subset(trainp.df, clusterTrain1 == 2)
proteintrain3 <- subset(trainp.df, clusterTrain1 == 3)

proteintrain21 <- subset(trainc.df, clusterTrain2 == 1)
proteintrain22 <- subset(trainc.df, clusterTrain2 == 2)
proteintrain23 <- subset(trainc.df, clusterTrain2 == 3)


proteintrain31 <- subset(trainf.df, clusterTrain3 == 1)
proteintrain32 <- subset(trainf.df, clusterTrain3 == 2)
proteintrain33 <- subset(trainf.df, clusterTrain3 == 3)


proteintrain41 <- subset(trains.df, clusterTrain4 == 1)
proteintrain42 <- subset(trains.df, clusterTrain4 == 2)
proteintrain43 <- subset(trains.df, clusterTrain4 == 3)




proteintest1 <- subset(testp.df, clusterTest1 == 1)
proteintest2 <- subset(testp.df, clusterTest1 == 2)
proteintest3 <- subset(testp.df, clusterTest1 == 3)

proteintest21 <- subset(testc.df, clusterTest2 == 1)
proteintest22 <- subset(testc.df, clusterTest2 == 2)
proteintest23 <- subset(testc.df, clusterTest2 == 3)

proteintest31 <- subset(testf.df, clusterTest3 == 1)
proteintest32 <- subset(testf.df, clusterTest3 == 2)
proteintest33 <- subset(testf.df, clusterTest3 == 3)


proteintest41 <- subset(tests.df, clusterTest4 == 1)
proteintest42 <- subset(tests.df, clusterTest4 == 2)
proteintest43 <- subset(tests.df, clusterTest4 == 3)

#building regression
trainp1 <- lm(Confirmed ~., data = proteintrain1)
trainp2 <- lm(Confirmed ~., data = proteintrain2)
trainp3 <- lm(Confirmed ~., data = proteintrain3)

trainf1 <- lm(Confirmed ~., data = proteintrain31)
trainf2 <- lm(Confirmed ~., data = proteintrain32)
trainf3 <- lm(Confirmed ~., data = proteintrain33)

trains1 <- lm(Confirmed ~., data = proteintrain41)
trains3 <- lm(Confirmed ~., data = proteintrain43)



#prediction
predictTest1 <- predict(trainp1, newdata = proteintest1)
predictTest2 <- predict(trainp2, newdata = proteintest2)
predictTest3 <- predict(trainp3, newdata = proteintest3)


predictTest31 <- predict(trainf1, newdata = proteintest31)
predictTest32 <- predict(trainf2, newdata = proteintest32)
predictTest33 <- predict(trainf3, newdata = proteintest33)

predictTest41 <- predict(trains1, newdata = proteintest41)
predictTest43 <- predict(trains3, newdata = proteintest43)


#Evaluate the accuracy
table(proteintest1$Confirmed, predictTest1 >= 0.5)
table(proteintest2$Confirmed, predictTest2 >= 0.5)
table(proteintest3$Confirmed, predictTest3 >= 0.5)



table(proteintest31$Confirmed, predictTest31 >= 0.5)
table(proteintest32$Confirmed, predictTest32 >= 0.5)
table(proteintest33$Confirmed, predictTest33 >= 0.5)

table(proteintest41$Confirmed, predictTest41 >= 0.5)
table(proteintest43$Confirmed, predictTest43 >= 0.5)

#To compute the overall test-set accuracy of the cluster-then-predict approach, we can combine all the test-set predictions into a single vector and all the true outcomes into a single vector:
allPredictions_protein <- c(predictTest1, predictTest2, predictTest3)
allOutcomes_protein <- c(proteintest1$Confirmed, proteintest2$Confirmed, proteintest3$Confirmed)



allPredictions_fats <- c(predictTest31, predictTest32, predictTest33)
allOutcomes_fats <- c(proteintest31$Confirmed, proteintest32$Confirmed, proteintest33$Confirmed)

allPredictions_supply <- c(predictTest41, predictTest43)
allOutcomes_supply <- c(proteintest41$Confirmed, proteintest42$Confirmed, proteintest43$Confirmed)
#Evalute the overall accuracy
table(allOutcomes_protein, allPredictions_protein>= 0.5)


```
Conclusion
Connecting all the observations above, we can conclude that one of the reasons why India has a lower COVID infection rate could be related to the country's dietary choices. Even though there are thousands of other factors that can impact infection rate in either countries, this analysis still proves that adapting to a healthier diet is one of the most important steps we can take to help reduce the amount of COVID cases.